---
title: "Honda Clustering"
author: "Raymond David, Hyoju Kang, Angel Hsu"
date: "2023-11-21"
output:
  pdf_document: default
  html_document: default
---


```{r include=FALSE}
# Load Libraries
library(tidyverse)
```


# Clustering
```{r}
features <- read_csv("/Users/hyoju/Desktop/Honda Competition/Datasets/feature.csv")
merged <- read_csv("/Users/hyoju/Desktop/Honda Competition/Datasets/merged.csv")
```
```{r}
library(stats)

# Function to extract numeric values
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select columns for clustering
selected_columns <- data.frame(
  Acceleration = extract_numeric(features$Acceleration),
  TopSpeed = extract_numeric(features$TopSpeed),
  Range = extract_numeric(features$Range),
  Efficiency = extract_numeric(gsub("[^0-9.]", "", features$Efficiency)),
  FastChargeSpeed = extract_numeric(features$FastChargeSpeed)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
#print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```

```{r}
# Assuming kmeans_result is your kmeans clustering result
plot(normalized_data, col = kmeans_result$cluster, pch = 19)
points(kmeans_result$centers, col = 1:k, pch = 8, cex = 2)

```
```{r}
# K-means + One Hot Encoding
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select numerical columns for clustering
selected_columns <- data.frame(
  Acceleration = extract_numeric(features$Acceleration),
  TopSpeed = extract_numeric(features$TopSpeed),
  Range = extract_numeric(features$Range),
  Efficiency = extract_numeric(gsub("[^0-9.]", "", features$Efficiency)),
  FastChargeSpeed = extract_numeric(features$FastChargeSpeed)
)

# One-hot encode categorical column 'Drive'
drive_column <- model.matrix(~ Drive - 1, data = features)
selected_columns <- cbind(selected_columns, drive_column)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
#print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)

```

# Cluster using Merged Dataset

```{r}
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select columns for clustering
selected_columns <- data.frame(
  SalesCount = merged$`Sales Count`,
  Acceleration = extract_numeric(merged$`Acceleration (sec)`),
  TopSpeed = extract_numeric(merged$`TopSpeed (km/h)`),
  Range = extract_numeric(merged$`Range (km)`)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
# Convert data.frame to matrix as kmeans() requires a matrix input
kmeans_result <- kmeans(as.matrix(normalized_data), centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```

```{r}
# K-means + One-hot Encoding
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Manually encode 'BodyStyle' column
encoded_body_style <- model.matrix(~ `Body Style` - 1, data = merged)

# Select numerical columns for clustering
selected_columns <- cbind(
  encoded_body_style,
  Acceleration = extract_numeric(merged$`Acceleration (sec)`),
  TopSpeed = extract_numeric(merged$`TopSpeed (km/h)`)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
# print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```

<Overall Interpretation>

# US Vehicle Feature Dataset 

### Cluster1:
Vehicles exhibit well-balanced performance attributes, slightly leaning towards enhanced efficiency and fast charging capabilities. Although All-Wheel Drive stands prevalent, it does not assert exclusive dominance within this cluster.

### Cluster2:
In contrast, the second cluster encompasses vehicles with diverse performance traits, markedly emphasizing speed and range. The presence of various drive types underscores a broad spectrum of vehicles, each possessing distinct performance features

### Cluster3:
Lastly, the third cluster suggests vehicles with more modest performance attributes, demonstrating a slight inclination towards acceleration while exhibiting lower efficiency, moderately reduced top speed, and range. The cluster portrays a mixed representation of vehicles, with a notable focus on front and rear-wheel drives 


# Washington EV Population Dataset

### Cluster1:
The initial group predominantly comprises hatchback cars with sales figures that align with the average. These vehicles notably exhibit higher acceleration but relatively lower top speeds, highlighting a trade-off between these performance metrics

### Cluster2:
The second cluster presents a blend of SUVs, fewer hatchbacks, and sedans. Vehicles in this cluster generally showcase higher acceleration but demonstrate lower top speeds, indicating a distinctive category of cars encompassing diverse body styles yet sharing analogous performance attributes

### Cluster3:
The third cluster primarily consists of sedans boasting sales counts higher than the average. These vehicles display lower acceleration but boast higher top speeds, possibly targeting a market segment prioritizing velocity over rapid acceleration