---
title: "Honda Analysis"
author: "Raymond David"
date: "2023-11-12"
output:
  pdf_document: default
  html_document: default
---

# Finding Correlation between Vehicle Ownership In Washington Vs The Overall US Market.
### Data Pre-processing
```{r message=FALSE}
# Loading necessary libraries
library(tidyverse)
# Importing Datasets

# Electric Vehicle Population Dataset
electric_pop <- read_csv("Electric_Vehicle_Population_Data.csv")
# Alternative Fuel Vehicles
alt_fuel <-read_csv("Alternative Fuel Vehicles US.csv")
# Public Charging Stations
public_charging <- read_csv("alt_fuel_stations (Nov 10 2023) Public.csv")
# Private Charging Stations
private_charging <- read_csv("alt_fuel_stations (Nov 10 2023) Private.csv")
```


```{r}
head(electric_pop)
```
```{r}
(electric_pop_grouped <-electric_pop %>% group_by(County)%>%
  summarize(count = n()))
```
```{r}
# Select Relevant Columns From Public Charging Dataset
(public_charging_subset <- select(public_charging,'Station Name','Street Address','City','State','ZIP','EV Level1 EVSE Num','EV Level2 EVSE Num',
       'EV DC Fast Count','Latitude','Longitude','EV Connector Types','Facility Type','EV Workplace Charging'))
```

```{r}
# Select Relevant Columns From Private Charging Dataset
(private_charging_subset <- select(private_charging,'Station Name','Street Address','City','State','ZIP','EV Level1 EVSE Num','EV Level2 EVSE Num',
       'EV DC Fast Count','Latitude','Longitude','EV Connector Types','Facility Type','EV Workplace Charging'))
```
```{r}
# Inner Join Public Charging and Electric Vehicle Population Dataset 
public_charging_subset$ZIP <- as.integer(public_charging_subset$ZIP)

(combined_public_EVpop <- inner_join(public_charging_subset, electric_pop, by = c("ZIP" = "Postal Code")))

```
### Regression Based on ZIP
```{r}
# Group electric pop by zip, showing counts
(counts_per_postal <- electric_pop %>%
  group_by(`Postal Code`) %>%
  summarize(count = n()) %>% 
  arrange(desc(count)))
```
```{r}
# Group public charging pop by zip, showing counts
(public_charging_counts <- public_charging_subset %>%
  group_by(ZIP) %>%
  summarize(count = n()) %>% 
  arrange(desc(count)))
```
```{r}
# Join Both Tables
# public_charging_subset$ZIP <- as.integer(public_charging_subset$ZIP)

(combined_public_EVpop <- inner_join(public_charging_counts, counts_per_postal, by = c("ZIP" = "Postal Code")))
colnames(combined_public_EVpop) <- c("ZIP","ChargingCounts", "VehicleCounts")
```

```{r}
# Run regression on ZIP
model <- lm(combined_public_EVpop$VehicleCounts ~ combined_public_EVpop$ChargingCounts)
summary(model)
```
Note: P-value is 0.0871 . Slightly larger than 0.05. 

### Run Regression on County level
```{r}
# Electric pop dataset grouped by County
(electric_pop_grouped <-electric_pop %>% group_by(County)%>%
  summarize(count = n()))
```

```{r}
# Combine dataset to borrow County Columns
(public_charging_grouped <- left_join(public_charging_subset, electric_pop, by = c("ZIP" = "Postal Code")))
```
```{r}
A <- (public_charging_grouped %>% group_by(County)%>% 
  summarize(count = n()))
```

```{r}
(combined_public_EVpop_county <- inner_join(A,electric_pop_grouped,by = c("County" = "County")))
colnames(combined_public_EVpop_county) <- c("County","ChargingCounts", "VehicleCounts")
```

```{r}
model2 <- lm(combined_public_EVpop_county$VehicleCounts ~ combined_public_EVpop_county$ChargingCounts)
summary(model2)
```
Note: Doing a County-level analysis makes sense

### Regression on Charging Type

```{r}
# Find counts of each charging type per zip
counts_per_county <- public_charging_subset  %>%
  group_by(ZIP) %>%
  summarize(
    Level1_Count = sum(!is.na(`EV Level1 EVSE Num`)),
    Level2_Count = sum(!is.na(`EV Level2 EVSE Num`)),
    DC_Fast_Count = sum(!is.na(`EV DC Fast Count`))
  )
```

```{r}
# Inner join datset by zip with electric pop
(joined_dataset <- left_join(counts_per_county,electric_pop, by = c("ZIP" = "Postal Code")))
```

```{r}
# Group dataset by County

(result <- joined_dataset %>%
  group_by(County) %>%
   summarize(
    Level1_Count = sum(Level1_Count),
    Level2_Count = sum(Level2_Count),
    DC_Fast_Count = sum(DC_Fast_Count),
    count = n()
  ) %>%
  select(County, Level1_Count, Level2_Count, DC_Fast_Count) %>%
  drop_na())
```
```{r}
joined_dataset_county <- inner_join(result, electric_pop_grouped,by = c("County" = "County"))
```


```{r}
# Run Regression
model3 <- lm(count ~ Level1_Count +Level2_Count + DC_Fast_Count, data = joined_dataset_county)
summary(model3)
```
**Notes:** Here, we noticed a trend that Level 2 counts and DC_Fast_Count are also important. 

# Regression Based on Private Dataset
### Run Regression on County level

```{r}
# Initial Data Study
# Most Common type of charger in Private Charging : Level 2
(result <- private_charging_subset %>%
  summarize(
    Level1_Count = sum(!is.na(`EV Level1 EVSE Num`)),
    Level2_Count = sum(!is.na(`EV Level2 EVSE Num`)),
    DC_Fast_Count = sum(!is.na(`EV DC Fast Count`))))
```

```{r}
# Most Common type of charger in Public Charging : Level 2
(result <- public_charging_subset %>%
  summarize(
    Level1_Count = sum(!is.na(`EV Level1 EVSE Num`)),
    Level2_Count = sum(!is.na(`EV Level2 EVSE Num`)),
    DC_Fast_Count = sum(!is.na(`EV DC Fast Count`))))
```

```{r}
# Combine dataset to borrow County Columns
private_charging_subset <- private_charging_subset %>%
  mutate(ZIP = as.numeric(ZIP))
(private_charging_grouped <- left_join(private_charging_subset, electric_pop, by = c("ZIP" = "Postal Code")))
```

```{r}
B <- (private_charging_grouped %>% group_by(County)%>% 
  summarize(count = n()))
```

```{r}
(combined_private_EVpop_county <- inner_join(B,electric_pop_grouped,by = c("County" = "County")))
colnames(combined_private_EVpop_county) <- c("County","ChargingCounts", "VehicleCounts")
```

```{r}
model4 <- lm(VehicleCounts ~ ChargingCounts, data = combined_private_EVpop_county)
summary(model4)
```
### Regression based on Charging Types
```{r}
# Find counts of each charging type per zip
counts_per_county <- private_charging_subset  %>%
  group_by(ZIP) %>%
  summarize(
    Level1_Count = sum(!is.na(`EV Level1 EVSE Num`)),
    Level2_Count = sum(!is.na(`EV Level2 EVSE Num`)),
    DC_Fast_Count = sum(!is.na(`EV DC Fast Count`))
  )
```

```{r}
# Inner join datset by zip with electric pop
(joined_dataset <- left_join(counts_per_county,electric_pop, by = c("ZIP" = "Postal Code")))
```

```{r}
# Group dataset by County

(result <- joined_dataset %>%
  group_by(County) %>%
   summarize(
    Level1_Count = sum(Level1_Count),
    Level2_Count = sum(Level2_Count),
    DC_Fast_Count = sum(DC_Fast_Count),
    count = n()
  ) %>%
  select(County, Level1_Count, Level2_Count, DC_Fast_Count) %>%
  drop_na())
```

```{r}
joined_dataset_county <- inner_join(result, electric_pop_grouped,by = c("County" = "County"))

```


```{r}
# Run Regression
model5 <- lm(count ~ Level1_Count +Level2_Count + DC_Fast_Count, data = joined_dataset_county)
summary(model5)
```
**Note:** Level 2 is the only one significant 

# Further Analysis on the Charging Stations Datasets
```{r}
head(public_charging_subset)
```

```{r}
(summary_result <- public_charging_subset %>%
  group_by(`EV Connector Types`) %>%
  count() %>%
  distinct() %>% 
  arrange(desc(n)))
```


```{r}
# Improve on the code before using Regex
(summary_result <- public_charging_subset %>% 
 separate_rows(`EV Connector Types`, sep = " ") %>%
  filter(`EV Connector Types` %in% c("J1772", "CCS", "CHAdeMO", "NACS")) %>%
  group_by(`EV Connector Types`) %>%
  count() %>%
  arrange(desc(n)))
```
### Summary of Most Popular Charging Types from the Public Charging Dataset
```{r}
(summary_result <- public_charging_subset %>%
  separate_rows(`EV Connector Types`, sep = " ") %>%
  mutate(`EV Connector Types` = tolower(`EV Connector Types`)) %>%
  mutate(Type_Category = case_when(
    `EV Connector Types` %in% c("j1772", "ccs", "chademo", "nacs","tesla","nema1450","nema515","nema520","j1772combo") ~ `EV Connector Types`,
    TRUE ~ "Other"
  )) %>%
  group_by(Type_Category) %>%
  count() %>%
  arrange(desc(n)))
```


### Summary of Most Popular Charging Types from the Private Charging Dataset
```{r}
(summary_result <- private_charging_subset %>%
  separate_rows(`EV Connector Types`, sep = " ") %>%
  mutate(`EV Connector Types` = tolower(`EV Connector Types`)) %>%
  mutate(Type_Category = case_when(
    `EV Connector Types` %in% c("j1772", "ccs", "chademo", "nacs","tesla","nema1450","nema515","nema520","j1772combo") ~ `EV Connector Types`,
    TRUE ~ "Other"
  )) %>%
  group_by(Type_Category) %>%
  count() %>%
  arrange(desc(n)))
```

### Summary of Facility Types from Public Charging Dataset
```{r}
(summary_result <- public_charging_subset %>%
  group_by(`Facility Type`) %>%
  count() %>%
  distinct() %>% 
  arrange(desc(n))) %>% 
  drop_na()
```
### Summary of Facility Types from Private Charging Dataset
```{r}
(summary_result <- private_charging_subset %>%
  group_by(`Facility Type`) %>%
  count() %>%
  distinct() %>% 
  arrange(desc(n)))
```
### Summary of Availability of Workplace Charging from Public Charging Dataset
```{r}
(summary_result <- public_charging_subset %>%
  group_by(`EV Workplace Charging`) %>%
  count() %>%
  distinct() %>% 
  arrange(desc(n)))
```
### Summary of Availability of Workplace Charging from Private Charging Dataset
```{r}
(summary_result <- private_charging_subset %>%
  group_by(`EV Workplace Charging`) %>%
  count() %>%
  distinct() %>% 
  arrange(desc(n)))
```
